```{r setup, include = FALSE}
knitr::opts_chunk$set(message = F, 
                      error = F, 
                      warning = F)

```

# Week 8 {-} 
<div style = "font-size: 28pt"> **_Lotka-Volterra model of competition: linear stability analysis_**</div>

## Lecture in a nutshell {-}

* **Topic:**
    1. 
    2. 
    3. 

$\frac {dN_{A}}{dt} = N_{A}(r_{A}-\alpha_{AA}N_{A}-\alpha_{AB}N_{B})$
$\frac {dN_{B}}{dt} = N_{B}(r_{B}-\alpha_{BB}N_{B}-\alpha_{BA}N_{A})$

$J = \begin{vmatrix}\frac {\partial }{\partial N_{A}} \frac {dN_{A}}{dt} & 0\\0 & 0\end{vmatrix}$
$J = \begin{vmatrix}0 & 0\\0 & 0\end{vmatrix}$

    
<div style="height:1px ;"><br></div>

* **Topic:**
    1. 
    2. 
    3.

<div style="height:1px ;"><br></div>    
<br>


## Lab demonstration {-}

In this lab, we are going to analyze the two-species Lotka-Volterra competition model numerically and visualize the population dynamics under different parameter settings.

```{r, fig.width = 5, fig.height = 3.5, out.width = "70%", fig.align = "center", include=F, echo=F}
library(tidyverse)
library(deSolve)

LV_pop_dynamics <- function(r1 = 1.4, r2 = 1.2, a11 = 1/200, a21 = 1/400, a22 = 1/200, a12 = 1/300, N1_0 = 10, N2_0 = 10) {
    
### Model specification
LV_competition_model <- function(times, state, parms) {
  with(as.list(c(state, parms)), {
    dN1_dt = N1*(r1-a11*N1-a12*N2)  
    dN2_dt = N2*(r2-a22*N2-a21*N1)
    return(list(c(dN1_dt, dN2_dt)))
  })
}

### Model parameters
times <- seq(0, 100, by = 0.1)
state <- c(N1 = N1_0, N2 = N2_0)
parms <- c(r1 = r1, r2 = r2, a11 = a11, a21 = a21, a22 = a22, a12 = a12)

### Model application
pop_size <- ode(func = LV_competition_model, times = times, y = state, parms = parms)

### Visualize the population dynamics
pop_size %>%
  as.data.frame() %>%
  pivot_longer(cols = -time, names_to = "species", values_to = "N") %>%
  ggplot(aes(x = time, y = N, color = species)) + 
  geom_line(size = 1.5) +
  theme_classic(base_size = 12) +
  labs(x = "Time", y = "Population size") +
  scale_x_continuous(limits = c(0, 100.5), expand = c(0, 0)) +
  scale_y_continuous(limits = c(0, max(pop_size)*1.2), expand = c(0, 0)) +
  scale_color_brewer(name = NULL, palette = "Set1")
}

### Different parameter settings
LV_pop_dynamics(r1 = 1.4, r2 = 1.2, a11 = 1/200, a21 = 1/400, a22 = 1/200, a12 = 1/300, N1_0 = 10, N2_0 = 10)  # stable coexistence
LV_pop_dynamics(r1 = 1.2, r2 = 1.2, a11 = 1/200, a21 = 1/100, a22 = 1/100, a12 = 1/300, N1_0 = 10, N2_0 = 10)  # N1 wins
LV_pop_dynamics(r1 = 1.2, r2 = 1.2, a11 = 1/200, a21 = 1/500, a22 = 1/500, a12 = 1/300, N1_0 = 10, N2_0 = 10)  # N2 wins

```
<br>

## Additional readings {-}

<br>
<br>
<br>
<br>
<br>

## Assignments {-}

[Linear Stability Analysis of Lotka-Volterra Competition Model](./Assignments/Week8_LV Competition Model.pdf){target="_blank"}

<!-- [Suggested Solutions](./Assignments/Week8_LV Competition Model_with_Solutions.pdf){target="_blank"} -->





