```{r setup, include = FALSE}
knitr::opts_chunk$set(message = F, 
                      error = F, 
                      warning = F)

```

# Week 10 {-} 
<div style = "font-size: 28pt"> **_Modern coexistence theory & Lotka-Volterra model of predator-prey interactions (1)_**</div>

## Lecture in a nutshell {-}

* **Modern coexistence theory**
    1. From mutual invasion analysis for coexistence: $\frac{\alpha_{AB}}{\alpha_{BB}}<\frac{r_{A}}{r_{B}}<\frac{\alpha_{AA}}{\alpha_{BA}}$ 
    2. Multiplying all terms by $\sqrt{\frac{\alpha_{BB}\alpha_{BA}}{\alpha_{AA}\alpha_{AB}}}$: 
    $\sqrt{\frac{\alpha_{AB}\alpha_{BA}}{\alpha_{AA}\alpha_{BB}}}<\frac{r_{A}}{r_{B}}\sqrt{\frac{\alpha_{BB}\alpha_{BA}}{\alpha_{AA}\alpha_{AB}}}<\sqrt{\frac{\alpha_{AA}\alpha_{BB}}{\alpha_{AB}\alpha_{BA}}}$
        * $\sqrt{\frac{\alpha_{AB}\alpha_{BA}}{\alpha_{AA}\alpha_{BB}}}$: niche overlap ($\rho$) (niche difference: $1-\rho$)
        * $\frac{r_{A}}{r_{B}}$: demographic ratio
        * $\sqrt{\frac{\alpha_{BB}\alpha_{BA}}{\alpha_{AA}\alpha_{AB}}}$: sensitivity to competition
        * $\frac{r_{A}}{r_{B}}\sqrt{\frac{\alpha_{BB}\alpha_{BA}}{\alpha_{AA}\alpha_{AB}}}$: fitness ratio/competitive hierarchy ($\frac{K_{A}}{K_{B}}$)
    3. Coexistence: fitness ratio $\frac{K_{A}}{K_{B}}$ should be bounded between $\rho$ and $\frac{1}{\rho}$ 
    4. Competition outcomes:
    
```{r, fig.width = 5, fig.height = 5, out.width = "70%", fig.align = "center", echo = F}
library(tidyverse)

ggplot() + 
  geom_abline(slope = -1, )


```

<div style="height:1px ;"><br></div>

* **Lotka-Volterra model of predator-prey interactions**
    1. Model derivation:
<div style="margin-left: 30%;">$\begin{align}\frac {dN}{dt} = g(N)-f(N,P)\end{align}\\$</div>
<div style="margin-left: 30%; margin-bottom: 5px;">$\begin{align}\frac {dP}{dt} = ef(N,P)-d(P)\end{align}$</div>
        * $g(N)$: prey growth
        * $f(N,P)$: prey consumed by predator
        * $ef(N,P)$: prey conversion into predator
        * $d(P)$: predator mortality
    2. Functional response $f(N,P)$:
        * Type I: $f(N,P) = aN$
        * Type II: $f(N,P) = \frac{aN}{1+ahN}$ (_h_ is the handling time) or $\frac{uN}{K_{m}+N}$
        * Type III: $f(N,P) = \frac{aN^{2}}{1+ahN^{2}}$ or $\frac{uN^{2}}{K_{m}+N^{2}}$
    3. Various model forms:
        * Exponential growth of prey + Type I functional response
        * Logistic growth of prey + Type I functional response
        * Exponential growth of prey + Type II functional response
        * Logistic growth of prey + Type II functional response

<div style="height:1px ;"><br></div>
* **Model 1. Exponential growth of prey + Type I functional response**
    1. Model equations:
<div style="margin-left: 30%;">$\begin{align}\frac {dN}{dt} = rN-aNP\end{align}\\$</div>
<div style="margin-left: 30%; margin-bottom: 5px;">$\begin{align}\frac {dP}{dt} = eaNP-dP\end{align}$</div>
    2. The equilibrium points $(N^{*}, P^{*})$:
        * $E_{0} = (0, 0)$
        * $E_{NP} = (\frac{d}{ea}, \frac{r}{a})$
    3. Local stability analysis:
        1. $E_{0}$
            * $J_{E_{0}} = \begin{vmatrix}r & 0 \\ 0 & -d\end{vmatrix}$
            * $\lambda's = r$ and $-d$ (one positive and one negative)
            * Unstable
        2. $E_{NP}$
        * $J_{E_{NP}} = \begin{vmatrix}0 &  -a\frac{d}{ea} \\ ea\frac{r}{a} & 0\end{vmatrix}$
        * $\lambda's = \sqrt{rd}*i$ 
        * Real part is zero (pure imaginary number): the equilibrium is a center; sensitive to non-linear terms/initial conditions)
        * Lyapunov function (global stability analysis of the equilibrium) is useful for predicting the system dynamics in the situation where all real parts of the eigenvalues are zero 
        * In this case, the dynamics are neutral cycles

<div style="height:1px ;"><br></div>    
<br>

## Lab demonstration {-}

In this lab, we are going to analyze the classic Lotka-Volterra predator-prey model:

<div style="margin-left: 30%;">$\begin{align}\frac {dN}{dt} = rN-aNP\end{align}\\$</div>
<div style="margin-left: 30%; margin-bottom: 5px;">$\begin{align}\frac {dP}{dt} = eaNP-dP\end{align}$</div>

Please simulate the model using the parameter set (_N_ = 10, _P_ = 2, _r_ = 1.0, _a_ = 0.1, _e_ = 0.8, _d_ = 0.5) and plot the population trajectories of predator and prey over time. Also show their population dynamics in the state-space diagram.

What will happen if you add a perturbation to the system (i.e., change the initial conditions)? Try out different values of _N_ and _P_ and visualize the differences in the state-space diagram.

```{r, fig.width = 5, fig.height = 3.5, out.width = "70%", fig.align = "center"}
library(tidyverse)
library(deSolve)

### Model specification
LV_predation_model <- function(times, state, parms) {
  with(as.list(c(state, parms)), {
    dN_dt = r*N - a*N*P
    dP_dt = e*a*N*P - d*P
    return(list(c(dN_dt, dP_dt)))  
  })
}

### Model parameters
times <- seq(0, 1000, by = 0.005)
state <- c(N = 10, P = 2)
parms <- c(r = 1.0, a = 0.1, e = 0.8, d = 0.5)

### Model application
pop_size <- ode(func = LV_predation_model, times = times, y = state, parms = parms)

### Visualize the population dynamics
# (1) population trajectories
pop_size %>%
  as.data.frame() %>%
  pivot_longer(cols = -time, names_to = "species", values_to = "N") %>%
  ggplot(aes(x = time, y = N, color = species)) + 
  geom_line(size = 1.5) +
  theme_classic(base_size = 12) +
  labs(x = "Time", y = "Population size") +
  scale_x_continuous(limits = c(0, 100.5), expand = c(0, 0)) +
  scale_y_continuous(limits = c(0, max(pop_size[, -1])*1.2), expand = c(0, 0)) +
  scale_color_brewer(name = NULL, palette = "Set1", labels = c("Prey", "Predator"))

# (2) state-space diagram
pop_size %>%
  as.data.frame() %>%
  ggplot(aes(x = N, y = P)) + 
  geom_path() + 
  theme_classic(base_size = 12) +
  labs(x = "Prey", y = "Predator")

```
<br>

<style>
iframe {border: 0;}
</style>

<!-- Here is a shiny app for the Lotka-Volterra predator-prey model. Feel free to play around with different parameter inputs and see how the system dynamics change accordingly. -->

```{r, out.width = "800px", echo = F}
knitr::include_app("https://genchanghsu0115.shinyapps.io/LV_predation_mod_shinyapp/", height = "700px")
```

## Additional readings {-}

[Signs of stabilisation and stable coexistence](./Additional readings/Broekman_et_al_2019_Eco_Let.pdf){target="_blank"}
<br>

## Assignments {-}

[Stabilizing Lotka-Volterra Predator-Prey Model](./Assignments/Week10_LV Predation Model.pdf){target="_blank"}

<!-- [Suggested Solutions](./Assignments/Week10_LV Predation Model_with_Solutions.pdf){target="_blank"} -->
