```{r setup, include = FALSE}
knitr::opts_chunk$set(message = F, 
                      error = F, 
                      warning = F)

```

# Week 11 {-} 
<div style = "font-size: 28pt"> **_Lotka-Volterra model of predator-prey interactions (2)_**</div>

## Lecture in a nutshell {-}

* **Topic:**
    1. 
    2. 
    3. 
    
<div style="height:1px ;"><br></div>

* **Topic:**
    1. 
    2. 
    3.

<div style="height:1px ;"><br></div>    
<br>


## Lab demonstration {-}

Similar to what we've done in the previous class, in this lab we are going to analyze the Rosenzweig–MacArthur predator–prey model:

<div style="margin-left: 30%;">$\begin{align}\frac {dN}{dt} = rN(1-\frac{N}{K})-a\frac{N}{h+N}P\end{align}\\$</div>
<div style="margin-left: 30%; margin-bottom: 10px;">$\begin{align}\frac {dP}{dt} = ea\frac{N}{h+N}P-dP\end{align}$</div>

Please simulate the model using the parameter set (_N_ = 5, _P_ = 2, _r_ = 1.0, _K_ = 5.0, _a_ = 1.3, _h_ = 0.9, _e_ = 0.6, _d_ = 0.5) and plot the population trajectories of predator and prey as well as show their population dynamics in the state-space diagram.

What will happen if you add a perturbation to the system (i.e., change the initial conditions)? Try out different values of _N_ and _P_ and visualize the differences in the state-space diagram. Also compare the results of the Rosenzweig–MacArthur model and the original Lotka-Volterra model. What do you find regarding the final equilibrium cycles?

```{r, fig.width = 5, fig.height = 3.5, out.width = "70%", fig.align = "center", eval = F, include = F, echo = F}
library(tidyverse)
library(deSolve)

### Model specification
RW_predation_model <- function(times, state, parms) {
  with(as.list(c(state, parms)), {
    dN_dt = r*N*(1-(N/K))-(a*N/(1+a*h*N))*P
    dP_dt = e*(a*N/(1+a*h*N))*P-d*P
    return(list(c(dN_dt, dP_dt)))  
  })
}

### Model parameters
times <- seq(0, 200, by = 0.01)  
state <- c(N = 5, P = 2)  
parms <- c(r = 1.0, K = 5.0, a = 1.3, h = 0.9, e = 0.6, d = 0.5) 

### Model application
pop_size <- ode(func = RW_predation_model, times = times, y = state, parms = parms)

### Visualize the population dynamics
# (1) population trajectories
pop_size %>%
  as.data.frame() %>%
  pivot_longer(cols = -time, names_to = "species", values_to = "N") %>%
  ggplot(aes(x = time, y = N, color = species)) + 
  geom_line(size = 1.5) +
  theme_classic(base_size = 12) +
  labs(x = "Time", y = "Population size") +
  scale_x_continuous(limits = c(0, 200.5), expand = c(0, 0)) +
  scale_y_continuous(limits = c(0, max(pop_size[, -1])*1.2), expand = c(0, 0)) +
  scale_color_brewer(name = NULL, palette = "Set1", labels = c("Prey", "Predator"))

# (2) state-space diagram
pop_size %>%
  as.data.frame() %>%
  ggplot(aes(x = N, y = P)) + 
  geom_path() + 
  geom_vline(xintercept = with(as.list(parms), d/(e*a-a*d*h)), color = "#377EB8", size = 1) +
  geom_function(data = data.frame(x = seq(0, 5, 0.01)), aes(x = x), fun = function(n){with(as.list(parms), r*(1+a*h*n)*(1-n/K)/a)}, inherit.aes = F, color = "#E41A1C", size = 1) + 
  geom_point(aes(x = tail(N, 1), y = tail(P, 1)), size = 2) +
  theme_classic(base_size = 14) +
  theme(axis.line.x = element_line(color = "#377EB8", size = 1),
        axis.line.y = element_line(color = "#E41A1C", size = 1)) + 
  labs(x = "Prey", y = "Predator") + 
  scale_x_continuous(limits = c(0, max(pop_size[, "N"]*1.05)), expand = c(0, 0)) + 
  scale_y_continuous(limits = c(0, max(pop_size[, "P"]*1.05)), expand = c(0, 0)) 

```
<br>

<style>
iframe {border: 0;}
</style>

Here is a shiny app for the Rosenzweig–MacArthur predator–prey model. Feel free to play around with different parameter inputs and see how the system dynamics change accordingly.

```{r, out.width = "800px", echo = F}
knitr::include_app("https://genchanghsu0115.shinyapps.io/LV_predation_mod_shinyapp/", height = "700px")
```

### Special topic: time-scale separation {-}

Time-scale separation is a useful technique to reduce the dimension of the model system, where some state variables are assumed to be at a much shorter time scale (i.e., fast variables) compared with the others (i.e., slow variables). The fast variables will respond rapidly to a slight change in the slow variables so that the slow variables can be viewed as "constants" from the perspective of fast variables.

Here, we are going to use the Lotka-Volterra model with logistic prey growth to demonstrate time-scale separation: one way to achieve this is to increase the prey growth rate so that prey will grow much faster than predator, mimicking the shorter generation time of prey relative to that of predator.

If you change _r_, you can see that the population dynamics of prey and predator change quite dramatically. Some interesting patterns will arise!

<style>
iframe {border: 0;}
</style>

```{r, out.width = "800px", echo = F}
knitr::include_app("https://genchanghsu0115.shinyapps.io/LV_predation_mod_shinyapp/", height = "700px")
```

## Additional readings {-}

<br>
<br>
<br>
<br>
<br>

## Assignments {-}

[](./Assignments/.pdf){target="_blank"}


