```{r setup, include = FALSE}
knitr::opts_chunk$set(message = F, 
                      error = F, 
                      warning = F)

```

# Week 12 {-} 
<div style = "font-size: 28pt"> **_Resource competition: R* models_**</div>

## Lecture in a nutshell {-}

* **Topic:**
    1. 
    2. 
    3. 
    
<div style="height:1px ;"><br></div>

* **Topic:**
    1. 
    2. 
    3.

<div style="height:1px ;"><br></div>    
<br>


## Lab demonstration {-}

### Part 1 - R* rule {-}
In this part of the lab section, we are going to analyze the consumer-resource dynamics and demonstrate the R* rule:

<div style="margin-left: 30%;">$\begin{align}\frac {dN_{1}}{dt} = ea_{1}RN_{1}-dN_{1}\end{align}\\$</div>
<div style="margin-left: 30%;">$\begin{align}\frac {dN_{2}}{dt} = ea_{2}RN_{2}-dN_{2}\end{align}\\$</div>
<div style="margin-left: 30%; margin-bottom: 15px;">$\begin{align}\frac {dR}{dt} = d(S_{0}-R)-a_{1}RN_{1}-a_{2}RN_{2}\end{align}$</div>

Please simulate the model using the parameter set (_N1_ = 2, _N2_ = 2, _R_ = 0.1, _a1_ = 0.4, _a2_ = 0.6, _e_ = 1, _d_ = 0.01, _S~0~_ = 0.1) and visualize the population trajectories of the two consumers as well as the resource. 

Below is a plot of resource level vs. consumers' per capita growth rate. What would you predict about the system dynamics? 

```{r, fig.width = 5, fig.height = 3.5, out.width = "70%", fig.align = "center"}
library(tidyverse)
library(deSolve)

### Parameters
a1 <- 0.4 
a2 <- 0.6
e <- 1
d <- 0.01 
S0 <- 0.1

### Resource level vs. consumers' per capita population growth
data.frame(R = seq(0, 0.1, 0.001)) %>%
  mutate(N1 = e*a1*R-d,
         N2 = e*a2*R-d) %>%
  pivot_longer(-R, names_to = "species", values_to = "N") %>%
  ggplot(aes(x = R, y = N, color = species)) +
  geom_line(size = 1.2) + 
  geom_hline(yintercept = 0, linetype = "dashed") + 
  labs(x = "Resource level", y = "Per capita growth rate") +
  scale_x_continuous(limits = c(0, 0.11), expand = c(0, 0)) +
  scale_y_continuous(limits = c(NA, 0.05), expand = c(0, 0)) +
  scale_color_brewer(name = NULL, palette = "Set1") + 
  theme_classic(base_size = 14)

```
```{r, fig.width = 6, fig.height = 3.5, out.width = "90%", fig.align = "center"}
### Model specification
CR_model <- function(times, state, parms){
  with(as.list(c(state, parms)), {
    dN1_dt = e*a1*R*N1 - d*N1
    dN2_dt = e*a2*R*N2 - d*N2
    dR_dt =  d*(S0-R) - a1*R*N1 - a2*R*N2
    return(list(c(dN1_dt, dN2_dt, dR_dt)))
  })
}

### Model parameters
times <- seq(0.1, 1000, by = 0.1)  
state <- c(N1 = 2, N2 = 2, R = 0.1)  
parms <- c(a1 = 0.4, a2 = 0.6, e = 1, d = 0.01, S0 = 0.1) 

### Model application
pop_size <- ode(func = CR_model, times = times, y = state, parms = parms)

### Visualize the population dynamics
pop_size %>%
  as.data.frame() %>%
  pivot_longer(cols = -time, names_to = "species", values_to = "N") %>%
  mutate(trophic = case_when(species %in% c("N1", "N2") ~ "Consumer",
                             TRUE ~ "Resource")) %>%
  ggplot(aes(x = time, y = N, color = species)) + 
  geom_line(size = 1.5) +
  facet_wrap(~ trophic, 
             ncol = 2, 
             scales = "free_y",
             strip.position = "left") +
  theme_classic(base_size = 14) +
  theme(strip.background = element_blank(),
        strip.placement = "outside",
        legend.position = "top",
        legend.title = element_blank(),
        plot.margin = margin(r = 5)) + 
  labs(x = "Time", y = NULL) +
  scale_x_continuous(limits = c(0, 1050), expand = c(0, 0)) +
  scale_y_continuous(limits = c(0, NA), expand = c(0, 0)) +
  scale_color_brewer(name = NULL, palette = "Set1")

```
<div style="height:15px ;"><br></div>

We can see that _N~1~_ goes extinct, as predicted by the R* rule; the remaining consumer _N~2~_ and the resource _R_ exhibit stable coexistence.

<br>

### Part 2 - Tilman’s resource ratio hypothesis {-}

```{r, fig.width = 5, fig.height = 3.5, out.width = "70%", fig.align = "center"}
library(tidyverse)
library(deSolve)

### Model specification
RM_predation_model <- function(times, state, parms) {
  with(as.list(c(state, parms)), {
    dN_dt = r*N*(1-(N/K))-(a*N/(1+a*h*N))*P
    dP_dt = e*(a*N/(1+a*h*N))*P-d*P
    return(list(c(dN_dt, dP_dt)))  
  })
}

### Model parameters
times <- seq(0, 200, by = 0.01)  
state <- c(N = 5, P = 2)  
parms <- c(r = 1.0, K = 5.0, a = 1.3, h = 0.9, e = 0.6, d = 0.5) 

### Model application
pop_size <- ode(func = RM_predation_model, times = times, y = state, parms = parms)

### Visualize the population dynamics
# (1) population trajectories
pop_size %>%
  as.data.frame() %>%
  pivot_longer(cols = -time, names_to = "species", values_to = "N") %>%
  ggplot(aes(x = time, y = N, color = species)) + 
  geom_line(size = 1.5) +
  theme_classic(base_size = 12) +
  labs(x = "Time", y = "Population size") +
  scale_x_continuous(limits = c(0, 200.5), expand = c(0, 0)) +
  scale_y_continuous(limits = c(0, max(pop_size[, -1])*1.2), expand = c(0, 0)) +
  scale_color_brewer(name = NULL, palette = "Set1", labels = c("Prey", "Predator"), direction = -1)

# (2) state-space diagram
pop_size %>%
  as.data.frame() %>%
  ggplot(aes(x = N, y = P)) + 
  geom_path() + 
  geom_vline(xintercept = with(as.list(parms), d/(e*a-a*d*h)), color = "#E41A1C", size = 1) +
  geom_function(data = data.frame(x = seq(0, 5, 0.01)), aes(x = x), fun = function(n){with(as.list(parms), r*(1+a*h*n)*(1-n/K)/a)}, inherit.aes = F, color = "#377EB8", size = 1) + 
  geom_point(aes(x = tail(N, 1), y = tail(P, 1)), size = 2) +
  theme_classic(base_size = 14) +
  theme(axis.line.x = element_line(color = "#E41A1C", size = 1),
        axis.line.y = element_line(color = "#377EB8", size = 1)) + 
  labs(x = "Prey", y = "Predator") + 
  scale_x_continuous(limits = c(0, max(pop_size[, "N"]*1.05)), expand = c(0, 0)) + 
  scale_y_continuous(limits = c(0, max(pop_size[, "P"]*1.05)), expand = c(0, 0)) 

```

<br>

### Part 3 - Relative nonlinearity {-}

```{r, fig.width = 5, fig.height = 3.5, out.width = "70%", fig.align = "center"}
library(tidyverse)
library(deSolve)

### Model specification
RM_predation_model <- function(times, state, parms) {
  with(as.list(c(state, parms)), {
    dN_dt = r*N*(1-(N/K))-(a*N/(1+a*h*N))*P
    dP_dt = e*(a*N/(1+a*h*N))*P-d*P
    return(list(c(dN_dt, dP_dt)))  
  })
}

### Model parameters
times <- seq(0, 200, by = 0.01)  
state <- c(N = 5, P = 2)  
parms <- c(r = 1.0, K = 5.0, a = 1.3, h = 0.9, e = 0.6, d = 0.5) 

### Model application
pop_size <- ode(func = RM_predation_model, times = times, y = state, parms = parms)

### Visualize the population dynamics
# (1) population trajectories
pop_size %>%
  as.data.frame() %>%
  pivot_longer(cols = -time, names_to = "species", values_to = "N") %>%
  ggplot(aes(x = time, y = N, color = species)) + 
  geom_line(size = 1.5) +
  theme_classic(base_size = 12) +
  labs(x = "Time", y = "Population size") +
  scale_x_continuous(limits = c(0, 200.5), expand = c(0, 0)) +
  scale_y_continuous(limits = c(0, max(pop_size[, -1])*1.2), expand = c(0, 0)) +
  scale_color_brewer(name = NULL, palette = "Set1", labels = c("Prey", "Predator"), direction = -1)

# (2) state-space diagram
pop_size %>%
  as.data.frame() %>%
  ggplot(aes(x = N, y = P)) + 
  geom_path() + 
  geom_vline(xintercept = with(as.list(parms), d/(e*a-a*d*h)), color = "#E41A1C", size = 1) +
  geom_function(data = data.frame(x = seq(0, 5, 0.01)), aes(x = x), fun = function(n){with(as.list(parms), r*(1+a*h*n)*(1-n/K)/a)}, inherit.aes = F, color = "#377EB8", size = 1) + 
  geom_point(aes(x = tail(N, 1), y = tail(P, 1)), size = 2) +
  theme_classic(base_size = 14) +
  theme(axis.line.x = element_line(color = "#E41A1C", size = 1),
        axis.line.y = element_line(color = "#377EB8", size = 1)) + 
  labs(x = "Prey", y = "Predator") + 
  scale_x_continuous(limits = c(0, max(pop_size[, "N"]*1.05)), expand = c(0, 0)) + 
  scale_y_continuous(limits = c(0, max(pop_size[, "P"]*1.05)), expand = c(0, 0)) 

```







## Additional readings {-}

[Competitive Exclusion](./Additional readings/Armstrong_&_McGehee_1980_AmNat.pdf){target="_blank"}

[A Graphical-Mechanistic Approach to Competition and Predation](./Additional readings/Tilman_1980_AmNat.pdf){target="_blank"}

<br>

## Assignments {-}

[Apparent Competition and P* Rule](./Assignments/Week12_Apparent Competition.pdf){target="_blank"}


