```{r setup, include = FALSE}
knitr::opts_chunk$set(message = F, 
                      error = F, 
                      warning = F)

```

# Week 12 {-} 
<div style = "font-size: 28pt"> **_Resource competition: R* models_**</div>

## Lecture in a nutshell {-}

* **Topic:**
    1. 
    2. 
    3. 
    
<div style="height:1px ;"><br></div>

* **Topic:**
    1. 
    2. 
    3.

<div style="height:1px ;"><br></div>    
<br>


## Lab demonstration {-}

### Part 1 - R* rule {-}
In this part of the lab section, we are going to analyze the consumer-resource dynamics with two consumers and one resource:

<div style="margin-left: 30%;">$\begin{align}\frac {dN_{1}}{dt} = ea_{1}RN_{1}-dN_{1}\end{align}\\$</div>
<div style="margin-left: 30%;">$\begin{align}\frac {dN_{2}}{dt} = ea_{2}RN_{2}-dN_{2}\end{align}\\$</div>
<div style="margin-left: 30%; margin-bottom: 15px;">$\begin{align}\frac {dR}{dt} = d(S_{0}-R)-a_{1}RN_{1}-a_{2}RN_{2}\end{align}$</div>

Please simulate the model using the parameter set (_N1_ = 2, _N2_ = 2, _R_ = 0.1, _a1_ = 0.4, _a2_ = 0.6, _e_ = 1, _d_ = 0.01, _S~0~_ = 0.1) and visualize the population trajectories of the two consumers as well as the resource. 

Below is a plot of resource level vs. consumers' per capita growth rate. What would you predict about the system dynamics? 

```{r, fig.width = 5, fig.height = 3.5, out.width = "70%", fig.align = "center"}
library(tidyverse)
library(deSolve)

### Parameters
a1 <- 0.4 
a2 <- 0.6
e <- 1
d <- 0.01 
S0 <- 0.1

### Resource level vs. consumers' per capita population growth
data.frame(R = seq(0, 0.1, 0.001)) %>%
  mutate(N1 = e*a1*R-d,
         N2 = e*a2*R-d) %>%
  pivot_longer(-R, names_to = "species", values_to = "N") %>%
  ggplot(aes(x = R, y = N, color = species)) +
  geom_line(size = 1.2) + 
  geom_hline(yintercept = 0, linetype = "dashed") + 
  labs(x = "Resource level", y = "Per capita growth rate") +
  scale_x_continuous(limits = c(0, 0.11), expand = c(0, 0)) +
  scale_y_continuous(limits = c(NA, 0.05), expand = c(0, 0)) +
  scale_color_brewer(name = NULL, palette = "Set1") + 
  theme_classic(base_size = 14)

```

```{r, fig.width = 7, fig.height = 3.5, out.width = "90%", fig.align = "center"}
### Model specification
CR_model_2C_1R <- function(times, state, parms){
  with(as.list(c(state, parms)), {
    dN1_dt = e*a1*R*N1 - d*N1
    dN2_dt = e*a2*R*N2 - d*N2
    dR_dt =  d*(S0-R) - a1*R*N1 - a2*R*N2
    return(list(c(dN1_dt, dN2_dt, dR_dt)))
  })
}

### Model parameters
times <- seq(0.1, 1000, by = 0.1)  
state <- c(N1 = 2, N2 = 2, R = 0.1)  
parms <- c(a1 = 0.4, a2 = 0.6, e = 1, d = 0.01, S0 = 0.1)  # R and S0 should be the same in the chemostat

### Model application
pop_size <- ode(func = CR_model_2C_1R, times = times, y = state, parms = parms)

### Visualize the population dynamics
pop_size %>%
  as.data.frame() %>%
  pivot_longer(cols = -time, names_to = "species", values_to = "N") %>%
  mutate(trophic = case_when(species %in% c("N1", "N2") ~ "Consumer",
                             TRUE ~ "Resource")) %>%
  ggplot(aes(x = time, y = N, color = species)) + 
  geom_line(size = 1.5) +
  facet_wrap(~ trophic, 
             ncol = 2, 
             scales = "free_y",
             strip.position = "left") +
  theme_classic(base_size = 14) +
  theme(strip.background = element_blank(),
        strip.placement = "outside",
        legend.position = "top",
        legend.title = element_blank(),
        plot.margin = margin(r = 5)) + 
  labs(x = "Time", y = NULL) +
  scale_x_continuous(limits = c(0, 1050), expand = c(0, 0)) +
  scale_y_continuous(limits = c(0, NA), expand = c(0, 0)) +
  scale_color_brewer(name = NULL, palette = "Set1")

```
<div style="height:15px ;"><br></div>

We can see that _N~1~_ goes extinct, as predicted by the R* rule; the remaining consumer _N~2~_ and the resource _R_ exhibit stable coexistence.

<br>

### Part 2 - Tilman’s resource ratio hypothesis {-}

Now we will extend the model in Part 1. into two consumers and two resources:

<div style="margin-left: 30%;">$\begin{align}\frac {dN_{1}}{dt} = ea_{1a}R_{a}N_{1}+ea_{1b}R_{b}N_{1}-dN_{1}\end{align}\\$</div>
<div style="margin-left: 30%;">$\begin{align}\frac {dN_{2}}{dt} = ea_{2a}R_{a}N_{2}+ea_{2b}R_{b}N_{2}-dN_{2}\end{align}\\$</div>
<div style="margin-left: 30%; margin-bottom: 15px;">$\begin{align}\frac {dR_{a}}{dt} = d(S_{0a}-R_{a})-a_{1a}R_{a}N_{1}-a_{2a}R_{a}N_{2}\end{align}$</div>
<div style="margin-left: 30%; margin-bottom: 15px;">$\begin{align}\frac {dR_{b}}{dt} = d(S_{0b}-R_{b})-a_{1b}R_{b}N_{1}-a_{2b}R_{b}N_{2}\end{align}$</div>

Here, the two consumers _N~1~_ and _N~2~_ compete for two substitutable resources _R~1~_ and _R~2~_. Under certain conditions, the two consumers are able to coexist:

```{r, fig.width = 7, fig.height = 3.5, out.width = "90%", fig.align = "center"}
library(tidyverse)
library(deSolve)

### Model specification
CR_model_2C_2R <- function(time, state, parms){
  with(as.list(c(state, parms)), {
    dN1_dt = e*a1a*Ra*N1 + e*a1b*Rb*N1 - d*N1
    dN2_dt = e*a2a*Ra*N2 + e*a2b*Rb*N2 - d*N2
    dRa_dt = d*(S0_a-Ra) - (a1a*N1*Ra) - (a2a*N2*Ra) 
    dRb_dt = d*(S0_b-Rb) - (a1b*N1*Rb) - (a2b*N2*Rb)
    return(list(c(dN1_dt, dN2_dt, dRa_dt, dRb_dt)))
  })
}

### Model parameters
times <- seq(0.1, 2000, by = 0.01)
state <- c(N1 = 0.05, N2 = 0.05, Ra = 0.3, Rb = 0.3)
parms <- c(a1a = 0.4, a1b = 0.8, a2a = 0.6, a2b = 0.5, e = 1, d = 0.1, S0_a = 0.3, S0_b = 0.3)  # Ra/Rb and S0a/S0b should be the same in the chemostat

### Model application
pop_size <- ode(func = CR_model_2C_2R, times = times, y = state, parms = parms)

### Visualize the population dynamics
pop_size %>%
  as.data.frame() %>%
  pivot_longer(cols = -time, names_to = "species", values_to = "N") %>%
  mutate(trophic = case_when(species %in% c("N1", "N2") ~ "Consumer",
                             TRUE ~ "Resource")) %>%
  ggplot(aes(x = time, y = N, color = species)) + 
  geom_line(size = 1.5) +
  facet_wrap(~ trophic, 
             ncol = 2, 
             scales = "free_y",
             strip.position = "left") +
  theme_classic(base_size = 14) +
  theme(strip.background = element_blank(),
        strip.placement = "outside",
        legend.position = "top",
        legend.title = element_blank(),
        plot.margin = margin(r = 8)) + 
  labs(x = "Time", y = NULL) +
  scale_x_continuous(limits = c(0, 2050), expand = c(0, 0)) +
  scale_y_continuous(limits = c(0, NA), expand = c(0, 0)) +
  scale_color_brewer(name = NULL, palette = "Set1")

```
<br>

The phase diagram also illustrates stable coexistence of the two consumers: 
```{r, fig.width = 5, fig.height = 3.5, out.width = "70%", fig.align = "center"}
### Parameters
a1a <- 0.4
a1b <- 0.8
a2a <- 0.6
a2b <- 0.5
e <- 1 
d <- 0.1 
S0_a <- 0.3
S0_b <- 0.3

### Slopes and intercepts of the ZNGI's
ZNGI_slope_N1 <- -a1a/a1b
ZNGI_intercept_N1 <- d/(e*a1b)  
ZNGI_slope_N2 <- -a2a/a2b
ZNGI_intercept_N2 <- d/(e*a2b)

### Consumption vectors
eqilibrium_Ra <- (d/e)*((a1b-a2b)/(a2a*a1b-a2b*a1a))
eqilibrium_Rb <- (d/e)*((a1a-a2a)/(a2b*a1a-a2a*a1b))

convec_df <- data.frame(x = c(eqilibrium_Ra + 6*a1a*eqilibrium_Ra, 
                              eqilibrium_Ra + 6*a2a*eqilibrium_Ra),
                        y = c(eqilibrium_Rb + 6*a1b*eqilibrium_Rb, 
                              eqilibrium_Rb + 6*a2b*eqilibrium_Rb),
                        xend = c(eqilibrium_Ra - a1a*eqilibrium_Ra, 
                              eqilibrium_Ra - a2a*eqilibrium_Ra),
                        yend = c(eqilibrium_Rb - a1b*eqilibrium_Rb, 
                              eqilibrium_Rb - a2b*eqilibrium_Rb),
                        species = c("N1", "N2"))

### Phase diagram
ggplot() + 
  geom_abline(slope = ZNGI_slope_N1, intercept = ZNGI_intercept_N1, color = "#E41A1C", size = 1.2) + 
  geom_abline(slope = ZNGI_slope_N2, intercept = ZNGI_intercept_N2, color = "#377EB8", size = 1.2) + 
  geom_segment(data = convec_df, aes(x = x, y = y, xend = xend, yend = yend, color = species), linetype = "blank") + 
  geom_segment(data = convec_df, aes(x = x, y = y, xend = xend, yend = yend, color = species), size = 0.5, linetype = "dashed", arrow = arrow(type = "closed", length = unit(0.1, "inches")), show.legend = F) +
  geom_path(data = as.data.frame(pop_size), aes(x = Ra, y = Rb)) +
  geom_point(data = as.data.frame(pop_size), aes(x = last(Ra), y = last(Rb)), size = 2.5) +
  theme_classic(base_size = 14) +
  labs(x = expression(italic(R[a])), y = expression(italic(R[b]))) +
  scale_x_continuous(limits = c(0, 0.5), expand = c(0, 0)) +
  scale_y_continuous(limits = c(0, 0.5), expand = c(0, 0)) +
  scale_color_brewer(name = NULL, palette = "Set1", 
                     guide = guide_legend(override.aes = list(
                       linetype = "solid", size = 1.2))) +
  coord_fixed(ratio = 1)


```

In this particular case, if the initial resource levels _R~a0~_ and _R~b0~_ fall within the upper area bounded by the two consumption vectors, then the two species will coexist.

### Part 3 - Relative nonlinearity {-}

```{r, fig.width = 5, fig.height = 3.5, out.width = "70%", fig.align = "center", eval=F, include=F, echo=F}
library(tidyverse)
library(deSolve)

### Model specification
RM_predation_model <- function(times, state, parms) {
  with(as.list(c(state, parms)), {
    dN_dt = r*N*(1-(N/K))-(a*N/(1+a*h*N))*P
    dP_dt = e*(a*N/(1+a*h*N))*P-d*P
    return(list(c(dN_dt, dP_dt)))  
  })
}

### Model parameters
times <- seq(0, 200, by = 0.01)  
state <- c(N = 5, P = 2)  
parms <- c(r = 1.0, K = 5.0, a = 1.3, h = 0.9, e = 0.6, d = 0.5) 

### Model application
pop_size <- ode(func = RM_predation_model, times = times, y = state, parms = parms)

### Visualize the population dynamics
# (1) population trajectories
pop_size %>%
  as.data.frame() %>%
  pivot_longer(cols = -time, names_to = "species", values_to = "N") %>%
  ggplot(aes(x = time, y = N, color = species)) + 
  geom_line(size = 1.5) +
  theme_classic(base_size = 12) +
  labs(x = "Time", y = "Population size") +
  scale_x_continuous(limits = c(0, 200.5), expand = c(0, 0)) +
  scale_y_continuous(limits = c(0, max(pop_size[, -1])*1.2), expand = c(0, 0)) +
  scale_color_brewer(name = NULL, palette = "Set1", labels = c("Prey", "Predator"), direction = -1)

# (2) state-space diagram
pop_size %>%
  as.data.frame() %>%
  ggplot(aes(x = N, y = P)) + 
  geom_path() + 
  geom_vline(xintercept = with(as.list(parms), d/(e*a-a*d*h)), color = "#E41A1C", size = 1) +
  geom_function(data = data.frame(x = seq(0, 5, 0.01)), aes(x = x), fun = function(n){with(as.list(parms), r*(1+a*h*n)*(1-n/K)/a)}, inherit.aes = F, color = "#377EB8", size = 1) + 
  geom_point(aes(x = tail(N, 1), y = tail(P, 1)), size = 2) +
  theme_classic(base_size = 14) +
  theme(axis.line.x = element_line(color = "#E41A1C", size = 1),
        axis.line.y = element_line(color = "#377EB8", size = 1)) + 
  labs(x = "Prey", y = "Predator") + 
  scale_x_continuous(limits = c(0, max(pop_size[, "N"]*1.05)), expand = c(0, 0)) + 
  scale_y_continuous(limits = c(0, max(pop_size[, "P"]*1.05)), expand = c(0, 0)) 

```

## Additional readings {-}

[Competitive Exclusion](./Additional readings/Armstrong_&_McGehee_1980_AmNat.pdf){target="_blank"}

[A Graphical-Mechanistic Approach to Competition and Predation](./Additional readings/Tilman_1980_AmNat.pdf){target="_blank"}

<br>

## Assignments {-}

[Apparent Competition and P* Rule](./Assignments/Week12_Apparent Competition.pdf){target="_blank"}

<!-- [Suggested Solutions](./Assignments/Week12_Apparent Competition_with_Solutions.pdf){target="_blank"} -->

