```{r setup, include = FALSE}
knitr::opts_chunk$set(message = F, 
                      error = F, 
                      warning = F)

```

# Week 13 {-} 
<div style = "font-size: 28pt"> **_Multispecies models of competition: apparent competition_**</div>

## Lecture in a nutshell {-}

* **Topic:**
    1. 
    2. 
    3. 
    
<div style="height:1px ;"><br></div>

* **Topic:**
    1. 
    2. 
    3.

<div style="height:1px ;"><br></div>    
<br>


## Lab demonstration {-}

Here, we are going to visualize the outcomes of apparent competition between two logistically-growing prey species _N~1~_ and _N~2~_ under linear predator consumption in the _r~1~/a~1~ - r~2~/a~2~_ space:

<div style="margin-left: 30%;">$\begin{align}\frac {dN_{1}}{dt} = r_{1}N_{1}(1-\frac{N_{1}}{K_{1}})-a_{1}N_{1}P\end{align}\\$</div>
<div style="margin-left: 30%;">$\begin{align}\frac {dN_{2}}{dt} = r_{2}N_{2}(1-\frac{N_{2}}{K_{2}})-a_{2}N_{2}P\end{align}\\$</div>
<div style="margin-left: 30%; margin-bottom: 15px;">$\begin{align}\frac {dP}{dt} = e_{1}a_{1}N_{1}P+e_{2}a_{2}N_{2}P-mp\end{align}$</div>

```{r, fig.width = 4, fig.height = 4, out.width = "75%", fig.align = "center", results = "hide"}
library(tidyverse)
library(deSolve)

### Model specification
Apparent_logistic_func <- function(r1, r2){
  Apparent_logistic_model <- function(times, state, parms) {
    with(as.list(c(state, parms)), {
      dN1_dt = r1*N1*(1-N1/K1) - a1*N1*P
      dN2_dt = r2*N2*(1-N2/K2) - a2*N2*P
      dP_dt = e1*a1*N1*P + e2*a2*N2*P - m*P
      return(list(c(dN1_dt, dN2_dt, dP_dt)))  
    })
  }
  
  ifelse(r1 < 0.2 & r2 < 0.2, 
         times <- seq(0, 1000, by = 1),  # longer simulation time for values near the origin
         times <- seq(0, 500, by = 1))
  
  state <- c(N1 = 5, N2 = 5, P = 1)
  parms <- c(r1 = r1, r2 = r2, a1 = 0.35, a2 = 0.35, e1 = 0.4, e2 = 0.5, m = 0.8, K1 = 15, K2 = 12)
  
  pop_size <- ode(func = Apparent_logistic_model, times = times, y = state, parms = parms)
  
  # get the final population size of N1 and N2
  N_final <- nrow(pop_size)
  N1_final <- pop_size[N_final, 2]
  N2_final <- pop_size[N_final, 3]
  
  # determine the outcome 
  if (N1_final < 1e-5) {
    return("N2_win")
  } else if (N2_final < 1e-5) {
    return("N1_win")
  } else {
    return("Coexist")
  }
  
}

### Apply the function
r_min <- 0
r_max <- 2.5
n <- 300

start_time <- Sys.time()  # the starting time of the simulations

comp_out <- expand.grid(r1 = seq(r_min, r_max, length.out = n),
                        r2 = seq(r_min, r_max, length.out = n)) %>%
  mutate(Outcome = map2_chr(r1, r2, Apparent_logistic_func))

end_time <- Sys.time()  # the ending time of the simulations

```

```{r}
end_time - start_time  # it might take a while!
```

```{r, fig.width = 4, fig.height = 4, out.width = "75%", fig.align = "center"}
### Visualize in the grids  
ggplot(comp_out, aes(xmin = r1/0.5, ymin = r2/0.5, 
                     xmax = r1/0.5 + (r_max-r_min)/(n-1)/0.5, ymax = r2/0.5 + (r_max-r_min)/(n-1)/0.5,
                     fill = Outcome)) + 
  geom_rect(show.legend = F, alpha = 0.9) + 
  scale_fill_brewer(palette = "Set1") + 
  scale_x_continuous(limits = c(r_min/0.5, (r_max+(r_max-r_min)/(n-1))/0.5 ), expand = c(0, 0)) +
  scale_y_continuous(limits = c(r_min/0.5, (r_max+(r_max-r_min)/(n-1))/0.5 ), expand = c(0, 0)) +
  labs(x = expression(r[1]/a[1]), y = expression(r[2]/a[2])) + 
  theme_bw(base_size = 14) + 
  theme(panel.grid = element_blank()) +
  coord_fixed(ratio = 1) + 
  annotate(geom = "text", x = 1, y = 4, label = "bold(bolditalic(N[2])~wins)", parse = T) + 
  annotate(geom = "text", x = 4, y = 4, label = "bold(Coexist)", parse = T) + annotate(geom = "text", x = 4, y = 1, label = "bold(bolditalic(N[1])~wins)", parse = T)

```

Note that the area of coexistence depends on the slopes of boundaries, which are a function of _m_, _e_, _a_, and _K_.

<br>

## Additional readings {-}

No additional readings this week.
<br>

## Assignments {-}

No assignments this week.


